---
title: "R_with_Ammar(Day_6)"
auther: "Waseem"
output:
  html_document:
    df_print: paged
  pdf_document: default
date: "2025-09-04"
---


## importing dataset

```{r}
df <- mtcars
head(df)
```
```{r}
tail(df)
```
## Import dataset
## making a heatmap
# we used as matrix because of to exclude categorical #variables 
# because meatmap only use numeric values
```{r}
# Heatmap
df <- as.matrix(df)
heatmap(df)
```

# lets normalize the value otherwise it will scatter plot # after normalization it will scale or measurable with #other
# all column are in scale 

```{r}
heatmap(df, scale = "column")
```
# lets add color cm.color is built-in function 
```{r}
heatmap(df, scale = "column", col = cm.colors(250))
```
```{r}
heatmap(df, scale = "column", col= terrain.colors(250))
```


# Important #
# use color blight free plot it will be helpful for #blight people

```{r}
library(RColorBrewer)
coul <- colorRampPalette(brewer.pal(8, "PiYG"))(25)
heatmap(df, scale="column", col = coul)
```


```{r}
library(RColorBrewer)
coul <- colorRampPalette(brewer.pal(8, "BrBG"))(25)
heatmap(df, scale="column", col = coul)
```

# click on settin to explore more about it
```{r R Col.Brewer}
display.brewer.all()
```

```{r}
volcano
```
# lets convert the data into wide formate from long #formate
```{r}
library(tidyverse)
volcano <- volcano %>% 
  
  # data wrangling , as_tibble to convert into tables
  as_tibble() %>%
  rowid_to_column(var= "X") %>%
  gather(key= "Y", value= "Z", -1) %>%
  
  # change Y to numeric
  mutate(Y=as.numeric(gsub("V", "", Y)))
```

# lets do plotting by piping operator .In piping no need #to write df
# for coloring we use fill(inner line) or color(outline) in R
```{r}
volcano %>% ggplot(aes(X, Y, col = Z))+geom_tile()

```
```{r}
volcano %>% ggplot(aes(X, Y, fill = Z))+geom_tile()
```




```{r}
library(tidyverse)

volcano_tbl <- as_tibble(volcano) %>%
  rowid_to_column(var = "x") %>%
  gather(key = "y", value = "z", -x) %>%
  mutate(y = as.numeric(gsub("V", "", y)))

ggplot(volcano_tbl, aes(x, y, fill = z)) +
  geom_tile()
```

```{r}
library(ggplot2)
library(hrbrthemes)
library(plotly)
library(dplyr)

# Dummy grid data
x <- LETTERS[1:20]
y <- paste0("var", 1:20)

# expand.grid creates all (X,Y) combinations
data <- expand.grid(X = x, Y = y)

# add random Z values
set.seed(123)  # reproducible
data$Z <- runif(nrow(data), 0, 5)

# add tooltip text
data <- data %>%
  mutate(text = paste0(
    "x: ", X, "\n",
    "y: ", Y, "\n",
    "value: ", round(Z, 2), "\n",
    "what else?"
  ))

head(data)

```









```{r}
# classic ggplot with tect in aes
p <- ggplot(data, aes(X, Y, fill= Z))+
  geom_tile()+
  theme_ipsum()
p
#pp <- ggplotly(p, tooltip= "text")
#pp
```
```{r}
# classic ggplot
ggplotly(p)
```
```{r}

```










